---
title: "Introduction to ipcwqrPIC package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ipcwqrPIC package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.width = 7,
  fig.height = 4,
  fig.align = "center",
  dpi = 150,
  fig.path = "vignettes/ipcwQR-"
)
```


## Introduction

`ipcwqrPIC` is the R package to fit a new inverse-probability censoring weighted (IPCW) estimating procedure for censored quantile regression when the data are partially interval-censored that include doubly-censored (DC) data and partly interval-censored (PIC) and possibly correlated within the same cluster.
Let $T$ and $X$ be the event time of interest and its related $p$-vector of covariates, respectively.
Our main objective is to estimate 
the $p$-dimensional quantile coefficient vector ${\boldsymbol{\beta}}_0(\tau)$
for some $\tau \in[\tau_L,\tau_R]\subset (0, 1)$ 
in the following linear quantile regression model:
$$
T_i = {\bf x}_i^T {\boldsymbol{\beta}}_0(\tau) + e_i(\tau),\quad i=1, \ldots ,n, 
$$
where $e_i(\tau)$ is the random error 
whose $\tau$th quantile conditional on ${\bf x}_i$ equals 0. 
When the data are subject to partially interval-censoring, 
left and right endpoints of the censoring time, $L$ and $R$,
are observed instead of $T$ such that $T\in(L,R)$.
Note that double-censoring  can also  be viewed as 
a special case of partly interval-censoring, 
i.e., $T$ is left-censored if $L=0$ and right-censored if $R=\infty$. 



## Usages

Installation of ipcwqrPIC package can be done by
```{r results='hide', message=FALSE, warning=FALSE}
devtools::install_github(repo="yejikim1202/ipcwqrPIC")
```

or
```{r results='hide', message=FALSE, warning=FALSE}
base::require("ipcwqrPIC")
```


picrq() function has the following arguments:
```{r eval=FALSE, message=FALSE, warning=FALSE}
picrq(U=U,V=V,delta=delta,x=x,tau=tau,estimation=NULL,var.estimation = "IS")
```

The option "estimation='DR'" is only applicable to DC data, so it should not be used for PIC data. Therefore, for PIC data, you should use "estimation=NULL" as the option.

dcrq() function has the following arguments:
```{r eval=FALSE, message=FALSE, warning=FALSE}
dcrq(L=L,R=R,T=T,delta=delta,x=x,tau=tau,var.estimation = "IS")
```

See the detailed description from help(picrq()) and help(dcrq()).

We first simulate univariate partly interval-censored (PIC) data with normal random error, which is similar to PIC data simulation setting of Kim et al. (2023+).

```{r message=FALSE, warning=FALSE}
library(ipcwqrPIC)
set.seed(111)
n = 200
x1 = runif(n,-1,1)
x2 = rbinom(n,1,0.43)
x = cbind(x1,x2)
T = 3 + x1 + x2 + rnorm(n)
U = (1 - 0.25*x1)*runif(n, -6, 5)
V = U + (1 - 0.1*x2)*runif(n, 6, 20)
U = exp(dplyr::case_when(TRUE ~ T, T>V ~ V, T<U ~ -Inf))
V = exp(dplyr::case_when(TRUE ~ T, T>V ~ Inf, T<U ~ U))
delta = ifelse(U==V, 1, 4)
tau=0.3
ipcwqrPIC::picrq(L=(U),R=(V),delta=delta,x=x,tau=tau,var.estimation = "IS")
```

We posit two estimating methods, ipcw estimating method and doubly robust ipcw estimating method, which can be conducted by specifying estimation = NULL and estimation = "DR", respectively.

Next, we simulate univariate doubly-censored (DC) data with normal random error, which is one of DC data simulation setting of Kim et al. (2023+).

```{r message=FALSE, warning=FALSE}
n=200; 
x1=runif(n,-1.2,1.7); x2=rbinom(n,1,0.6)
T = 7+x1+x2+rnorm(n)*(1-0.1*x2)
L=7+runif(n,-2.8,1.9); R=L+runif(n,4.2,8.1)
Y=pmin(R,pmax(T,L))
delta=case_when(
 T<L ~ 3,
 T>R ~ 2,
 TRUE ~ 1 #observed
)
L=L; R=R; T=T; delta=delta; x=cbind(x1,x2); tau=0.3
ipcwqrPIC::dcrq(L,R,T,delta,x,tau,var.estimation = "IS")
ipcwqrPIC::dcrq(L,R,T,delta,x,tau,var.estimation = "Bootstrap")
ipcwqrPIC::dcrq(L,R,T,delta,x,tau,estimation = "DR", var.estimation = "IS")
ipcwqrPIC::dcrq(L,R,T,delta,x,tau,wttype = "Beran", hlimit = 0.1, var.estimation = "IS")
ipcwqrPIC::dcrq(L,R,T,delta,x,tau,wttype = "Ishwaran", var.estimation = "IS")
```

We can choose one of three weighting types, the parametric KM estimating method, Beran's nonparametric KM estimating method, and Ishwaran's nonparametric survival random forests estimating method which can be accessible by using wttype="param", wttype="Beran" and wttype="Ishwaran", respectively.

Finally, we cover multivariate doubly-censored (MVDC) data with normal random error, which is one of MVDC data simulation setting of Kim et al. (2023+).

If id=NULL, we can gain the unadjusted IPCW method estimates, else if id=id and index value is put, we can get the adjusted IPCW method.
```{r message=FALSE, warning=FALSE}
cluster=60
v=rlnorm(cluster,mean=0,sd=0.5)
k = sample(0:9*10,cluster,replace = TRUE)
m = ifelse(quantile(v, k/100) <= v & quantile(v, (k+10)/100) <= v,
             k/10 + 2, 5)
id = as.numeric(unlist(apply(cbind(1:cluster,m),1,function(a) rep(a[1],a[2]))))
n=sum(m)
x1=runif(n,-1.4,0.9); x2=rbinom(n,1,0.5)
tabid=as.vector(table(id))
vi=as.vector(unlist(lapply(tabid, function(x) rnorm(x,0,0.5))))
T = 8+x1+x2 + vi +rnorm(n)
L=8+runif(n,-3.8,2); R=L+runif(n,3.2,9.7)
x=as.matrix(cbind(x1,x2),ncol=2)
delta=case_when(
    T<L ~ 3,
    T>R ~ 2,
    TRUE ~ 1 #observed
  )
ipcwqrPIC::dcrq(L=L,R=R,T=T,delta=delta,x=x,tau=tau,wttype="KM",var.estimation="IS",id=NULL)
ipcwqrPIC::dcrq(L=L,R=R,T=T,delta=delta,x=x,tau=tau,wttype="KM",var.estimation="IS",id=id,index = 1)
```

In real data analysis, we apply our proposed method to the multivariate clustered PIC data using a phase 3 metastatic colorectal cancer clinical trial. This dataset is accessible for data(mCRC) in the PICBayes R package (Pan, 2021).

Warning! If you use Beran's weighting, you should apply bootstrap estimation to variance estimating.

```{r message=FALSE, warning=FALSE}
library(PICBayes)
data("mCRC")
d = with(data.frame(mCRC), data.frame(U = ifelse(y==0,R,L),
                                      V = ifelse(y==2,L,R),
                                      # Cluster weighted data
                                      id=(rep(c(table(SITE)),c(table(SITE)))),
                                      # Treatment arm: 0 = FOLFIRI alone, 1 = Panitumumab + FOLFIRI.
                                      x1= case_when(TRT_C == 0 ~ 0, #Pan et al data
                                                    TRT_C == 1 ~ 1),
                                      # Tumor KRAS mutation status: 0 = wild-type, 1 = mutant.
                                      x2= case_when(KRAS_C == 0 ~ 1,
                                                    KRAS_C == 1 ~ 0),
                                      y = y
)); table(d$y)
delta=with(d,ifelse(y==3,1,
             ifelse(y==2,2,
                           ifelse(y==1,4,3)))); table(delta)
U=log(d$U);V=log(d$V); x = cbind(d$x1,d$x2); id=d$id;  tau=0.1;
ipcwqrPIC::picrq(L=U,R=V,delta=delta,x=x,tau=tau,wttype="Beran",hlimit=0.1,var.estimation = "Bootstrap",id=id,index = 1)
ipcwqrPIC::picrq(L=U,R=V,delta=delta,x=x,tau=tau,wttype="Ishwaran",hlimit=0.1,var.estimation = "Bootstrap",id=id,index = 1,B=100)
```


## References

* Beran, R. (1981). Nonparametric Regression with Randomly Censored Survival Data. Technical Report, Univ.California, Berkeley.

* Ishwaran, H., U. B. Kogalur, E. H. Blackstone, and M. S. Lauer (2008). Random survival forests. The annals of applied statistics. 2 (3), 841–860.

* Chiou, S. H., Kang, S. and Yan, J. (2015). 
Rank-based estimating equations with general weight for accelerated failure time models: an induced smoothing approach.
Statistics in Medicine 34(9): 1495–-1510.

* Gorfine, M., Y. Goldberg, and Y. Ritov (2017). A quantile regression model for failure-time data with time-
dependent covariates. Biostatistics. 18 (1), 132–146.

* Chiou, S. H., Kang, S. and Yan, J. (2015). Rank-based estimating equations with general weight for accelerated failure time models: an induced smoothing approach. Statistics in Medicine 34(9): 1495–-1510.

* Zeng, D. and D. Lin (2008). Efficient resampling methods for nonsmooth estimating functions. Biostatistics 9 (2), 355–363.

* Pan, C. (2021). 
PICBayes: Bayesian Models for Partly Interval-Censored Data. R package. 
https://CRAN.R-project.org/package=PICBayes.

* Kim, Y., Choi, T., Park, S., Choi, S. and Bandyopadhyay, D. (2023+). 
Inverse weighted quantile regression with partially interval-censored data.
*Submitted to SMMR*.